---
title: "data wrangling II"
author: "Kaila Boyer"
date: "9/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(haven)
```


## Import same data 
can list data directory in gitignore file to prevent it from going to github 

want to import 'FAS_litters.csv'

relative path (any situation) - 
```{r}
litters_df <- read_csv("data/FAS_litters.csv")
litters_df <- janitor::clean_names(litters_df)


pups_df <- read_csv("data/FAS_pups.csv")
pups_df <- janitor::clean_names(pups_df)

```

```{r}
select(litters_df, group, litter_number)
select(litters_df, group, gd0_weight, gd18_weight)

select(litters_df, group, gd0_weight:gd_of_birth)

select(litters_df, starts_with("pups"))

select(litters_df, GROUP = group, litter_number)

select(litters_df, litter_number, everything()) #rearranges columns 
relocate(litters_df, litter_number)
```

Can also use 'ends_with()' or 'contains()'
can rename variables in select(), group --> GROUP 

## Learning check 

In the pups data, select the columns containing litter number, sex, and PD ears.
```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## Filter 

```{r}
filter(litters_df, gd_of_birth == 20)
filter(litters_df, group == "Con7")

filter(litters_df, gd0_weight < 21)
filter(litters_df, gd0_weight > 23)

filter(litters_df, pups_survive != 4)
filter(litters_df, pups_survive == 4)

filter(litters_df, !(group == "Con7"))
filter(litters_df, group %in% c("Con7", "Con8")) #con7 or con8

filter(litters_df, group == "Con7" , gd_of_birth == 20)
filter(litters_df, group == "Con7" | gd_of_birth == 20)

drop_na(litters_df)
drop_na(litters_df, gd0_weight)
```
Double equal sign is a logical, single equal sign assigns value 

drop_na gets rid of all with missing values, adding a column to it removes rows with missing data for that variable 


## Mutate 

```{r}
mutate(litters_df, 
       weight_change = gd18_weight - gd0_weight, 
       group = str_to_lower(group))
```


## Arrange 

```{r}
arrange(litters_df, desc(gd0_weight))
arrange(litters_df, gd_of_birth, desc(gd0_weight))
```
Can arrange by multiple things 


## Pipes

```{r}
litters_data_raw <- read_csv("data/FAS_litters.csv")
litters_clean_name <- janitor::clean_names(litters_data_raw)
litters_select <- select(litters_clean_name, group,  pups_survive)
litters_filtered <- filter(litters_select, group == "Con7")

litters_df <- read_csv("data/FAS_litters.csv") %>% 
                  janitor::clean_names() %>% 
                  select(group, pups_survive) %>% 
                  filter(group == "Con7")


litters_df <- read_csv("data/FAS_litters.csv") %>% 
                  janitor::clean_names() %>% 
                  mutate(weight_change = gd18_weight - gd0_weight, 
                         group = str_to_lower(group)) %>% 
                  drop_na(weight_change) %>% 
                  filter(group %in% c("con7", "con8")) %>% 
                  select(litter_number, group, weight_change, everything())
```

# Learning Check 

Learning Assessment: Write a chain of commands that:

loads the pups data
cleans the variable names
filters the data to include only pups with sex 1
removes the PD ears variable
creates a variable that indicates whether PD pivot is 7 or more days


```{r}
read_csv("data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  filter(sex == 1) %>% 
  select(!pd_ears) %>% 
  mutate(PD_7 = pd_pivot >= 7)
```


