---
title: "data wrangling II"
author: "Kaila Boyer"
date: "9/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(haven)
```


## Import same data 
can list data directory in gitignore file to prevent it from going to github 

want to import 'FAS_litters.csv'

relative path (any situation) - 
```{r}
litters_df <- read_csv("data/FAS_litters.csv")
litters_df <- janitor::clean_names(litters_df)


pups_df <- read_csv("data/FAS_pups.csv")
pups_df <- janitor::clean_names(pups_df)

```

```{r}
select(litters_df, group, litter_number)
select(litters_df, group, gd0_weight, gd18_weight)

select(litters_df, group, gd0_weight:gd_of_birth)

select(litters_df, starts_with("pups"))

select(litters_df, GROUP = group, litter_number)

select(litters_df, litter_number, everything()) #rearranges columns 
relocate(litters_df, litter_number)
```

Can also use 'ends_with()' or 'contains()'
can rename variables in select(), group --> GROUP 

## Learning check 

In the pups data, select the columns containing litter number, sex, and PD ears.
```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## Filter 

```{r}
filter(litters_df, gd_of_birth == 20)
filter(litters_df, group == "Con7")

filter(litters_df, gd0_weight < 21)
filter(litters_df, gd0_weight > 23)

filter(litters_df, pups_survive != 4)
filter(litters_df, pups_survive == 4)

filter(litters_df, !(group == "Con7"))
filter(litters_df, group %in% c("Con7", "Con8")) #con7 or con8

filter(litters_df, group == "Con7" , gd_of_birth == 20)
filter(litters_df, group == "Con7" | gd_of_birth == 20)

drop_na(litters_df)
drop_na(litters_df, gd0_weight)

```
Double equal sign is a logical, single equal sign assigns value 

drop_na gets rid of all with missing values, adding a column to it removes rows with missing data for that variable 

