---
title: "Simple document"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

## Import same data 
can list data directory in gitignore file to prevent it from going to github 

want to import 'FAS_litters.csv'

relative path (any situation) - 
```{r}
litters_df <- read_csv("data/FAS_litters.csv")
```

absolute path (should not do) - cannot share with other people, gets messed up when you move the file 
```{r, eval = FALSE}
#read_csv("/Users/Owner/data_wrangling_i/data/FAS_litters.csv")
```

Clean variable names 
```{r}
names(litters_df)

litters_df <- janitor::clean_names(litters_df)
```

looking at the data 
```{r}
head(litters_df)
tail(litters_df)

skimr::skim(litters_df)
```
```{r, eval = FALSE}
view(litters_df)
```


## Arguments in `read_csv`
```{r}
litters_df <- read_csv(
  
  "data/FAS_litters.csv", 
  skip = 5, 
  col_names = FALSE, 
  na = "Low8"
)
```

## Parsing columns 

```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv",
  col_types = cols(
    Group = col_character(),
    `Litter Number` = col_character(),
    `GD0 weight` = col_double(),
    `GD18 weight` = col_double(),
    `GD of Birth` = col_integer(),
    `Pups born alive` = col_integer(),
    `Pups dead @ birth` = col_integer(),
    `Pups survive` = col_integer()
  )
)

tail(litters_data)
```


## Reading from excel 

reading MLB data 
```{r}
mlb11_df <- read_excel("data/mlb11.xlsx")

```
reading LOTR data 

```{r}
fellow_df <- read_excel("data/LotR_Words.xlsx", range = "B3:D6")
```

## Reading in SAS data 
tried to read in SAS file but says some features are unsupported
```{r, eval = FALSE}
pulse_df <- read_sas("data/public_pulse_data.sas7bdat")
```

## why to never use `read.csv`

it prints out the entire data table and provides no info about column types, row/col number like `read_csv`
it doesn't give `tibble`

```{r, eval = FALSE}
litter_df_base <- read.csv("data/FAS_litters.csv")
```


## Exporting data 

```{r}
write_csv(fellow_df, "data/fellowship_words.csv")
```





